
```{r setup, echo=FALSE, message=FALSE, warning=FALSE}
#Load packages

library(tidyverse)
library(GGally)
library(dplyr)
library(rpart)
library(rpart.plot)
library(caret)
library(ROSE)
library(ROCR)
library(readxl)
library(caret)
library(ggplot2)
library(neuralnet)
library(base)
library(nnet)
library(gridExtra)
library(NeuralNetTools)
library(Metrics)


# Loading Data
df <- read_excel("NNDataset.xlsx",sheet = "Sheet1")
df <- as.data.frame(df)
data <- df
```

Doing EDA for the Dataset from the case study.
Let's see what kind of data is present in the dataset. 
`r str(data)`

Let's check for null values in each of the columns in the data.
```{r}
#Check for null values in the data
colSums(is.na(data))

```

There are `r sum(is.na(data))` null values in the dataset.

From the case study, we see that the following variables are categorical: Fuel, MC, Auto, ABS,
Mfr_G,Abag_1,Abag_2, AC, Comp, CD, Clock, Pw, PStr, Radio, SpM, M_Rim, Tow_Bar.

Let's convert these variables into factors. But, before converting, let's check the unique values
in each of the columns.


```{r datacleaning}
# create a vector to store the names of columns with a unique value of 1
counts <- list()
cols_with_1_unique <- character()

for (col in names(data)) {
  count <- length(unique(data[, col]))
  counts[[col]] <- count
  if (count == 1){
    cols_with_1_unique <- c(cols_with_1_unique, col)
  }
}

cols_with_1_unique

```

We see that a few columns in the dataset have only one unique value. These variables would not be
helpful in training the neural net model. Therefore, we can remove these columns from our dataset.


```{r}
cleandata <- data[, !names(data) %in% cols_with_1_unique]
cd_lm <- cleandata

str(cleandata)

```

Visualizing the continuous variables in clean data.

```{r}
cont_variables <- c('Price', 'Age', 'KM','Wght')

cont_data <- cleandata[cont_variables]
cat_data <- cleandata %>%
            select(-all_of(cont_variables))

price_hist <- ggplot(cont_data, aes(x=Price)) + geom_histogram(binwidth = bw.nrd0(cleandata$Price))
age_hist <- ggplot(cont_data, aes(x=Age)) + geom_histogram(binwidth = bw.nrd0(cleandata$Age))
km_hist <- ggplot(cont_data, aes(x=KM)) + geom_histogram(binwidth = bw.nrd0(cleandata$KM))
wght_hist <- ggplot(cont_data, aes(x=Wght)) + geom_histogram(binwidth = bw.nrd0(cleandata$Wght))


PlotsList<- list(price_hist,age_hist,km_hist,wght_hist)
g <- grid.arrange(grobs = PlotsList, nrow = 2, ncol = 2)

```

Visualizing the categorical variables in clean data.

```{r warning=FALSE}
Cat_data_fac <- data.frame(lapply(cat_data, as.factor))

plots <- list() 
for (i in colnames(Cat_data_fac)){ 
  
  individual_plot <- Cat_data_fac %>% 
    group_by_(i) %>% 
    summarize(count = n()) %>% 
    ggplot(aes_string(x=i, y="count")) +
    geom_col(position="dodge") +
    geom_text(aes(label = count), vjust = 0, position = position_dodge(width=0.8))
 
plots[[i]] <- individual_plot 
} 

m <- marrangeGrob(plots,nrow = 2, ncol=5) 
```

From the bar graphs above, we can see that  for the variables 'Auto', 'Abag_2', 'Mfr_G', 'Comp',
'Clock', 'Radio', 'Pw' have very high frequency for one unique value and minuscule frequency for the
remaining unique values. These kind of variables won't be useful for us while training an NN model as
there is not enough data to understand the relationship between these variables and the target 
variable (Price).

```{r removeimbalVars}
cleandata <- cleandata[, !names(cleandata) %in% c('Auto', 'Abag_2', 'Mfr_G', 'Comp',
'Clock', 'Radio', 'Pw')]

```


```{r BiVar}
pairs(cleandata[,-6])
ggcorr(cleandata[,-6], label = TRUE, label_round = 1)

```

From the correlation matrix above, we see that the variables 'HP', 'CC', 'Grs', 'Wght', 'AC' and
'M_Rim' are highly correlated with 'Price'.


### (1) **After your EDA, what factors do you think influence a customer’s decision to buy a care? What are the objectives of the model that Farid plans to build?**

**(A)** After performing exploratory data analysis (EDA) in R on the dataset, it appears
that several factors may influence a customer's decision to buy a car. We see Horse power(HP), 
Cylinder volume in cubic centimeters(CC), Number of gear positions(Grs), Weight of the car(Wght),
Automatic Air conditioning(AC) and Metallic Rim(M_Rim) are the important variables which dictate
the price of the car. 

Let's discuss the role of eachof these variables on the price of the car.


1. **HP** - HP tell us whether the car is fast and is a major variable that can dictate the price.
2. **CC** - CC tells us the size of the engine. CC is strongly correlated with HP. Similar to HP, CC also dictates the price of the car.
3. **Grs** - Number of gears indicate the range of speed and control we can have on the car's performance. This var is also strongly correlated with both HP and CC and can dictate the price of the car.
4. **Wght** - Weight of the car tells us whether the car is a sedan, SUV etc. As the size of the car gets bigger, the price also increases.
5. **AC** - Automatic airconditioning is a feature that dictates the price as this is considered a necessary feature in cars now a days.
6. **M_Rim** - Having a metallic rim tells us whether the car is a premium model or a basic one. Hence, this variable also strongly dictates the price of the car.


The objectives of the model that Farid plans to build are to accurately predict the price of a car
based on these factors. By building a model that can accurately predict the price of a car based on
these factors, Farid can help Adam Ebraham decide on the new car models that are manufactured.
Additionally, by identifying the most important factors that influence a customer's decision to 
buy a car, Farid can focus on improving these factors in order to increase sales.

### (2) **Construct a neural network model. Validate and interpret the model using a different number of hidden neurons.**

**(A)** To construct a neural network model in R, we can use the 'nnet' package. This package
provides functions for creating, training, and evaluating neural network models.

But before building the model, we have to normalize our data.

```{r normalization, message=FALSE, warning=FALSE}
normalize <- function(x) 
{
  (x-min(x))/(max(x)-min(x))
}

cd <- cleandata %>% mutate_if(is.numeric, normalize)

```

Now that the data is normalized,  let's split the data into train and test data.

```{r splitdata}
set.seed(1234)
ind <- sample(2,nrow(cd), replace = T, prob = c(0.7,0.3))
train <- cd[ind == 1, ]
test <- cd[ind == 2, ]

```

**Neural Network Model**

```{r nnmodel, result='hide'}
naive_nn <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, 
                 size = 12, decay = 0.001, maxit = 1000)
```

```{r}
plotnet(naive_nn)

train_pred = predict(naive_nn, train)
test_pred = predict(naive_nn, test)

Accuracy_train <-function(x) 
{
  1-sqrt(mean((train$Price - x[,1])^2))
}

Accuracy_test <-function(x) 
{
  1-sqrt(mean((test$Price - x[,1])^2))
}
```

We have created the neural net model called "**naive_nn**" with number of hidden neurons of 12 which is trained on the train data. 

The accuracy of the model on train data = `r Accuracy_train(train_pred)`
The accuracy of the model on test data = `r Accuracy_test(test_pred)`

```{r Optdecay, results='hide', message=FALSE}

# let's select optimal decay parameter

set.seed(999)
indx <- sample(2, nrow(train), replace = T, prob = c(0.5,0.5))
train2 <- train[indx == 1, ]
valid <- train[indx == 2, ]

err <- vector("numeric", 100)
d <- seq(0.0001, 1, length.out = 100)
k = 1
for(i in d){
  nn_optdecay <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, 
                      size = 10, decay = i, maxit = 1000)
  pred.class <- predict(nn_optdecay, newdata = valid)
  err[k] <- sqrt(mean((pred.class-valid[,1])^2))
  k <- k+1
}

plot(d, err)
min_index <- which.min(err)
d_opt <- d[min_index]


```

We can use the optimal decay parameter found above to create neural network models with different hidden neurons and decide on the best model.


```{r Bestneuralnet, results='hide', message=FALSE}

#Neural network model with 2 hidden neurons
nn_2 <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size = 2,
                decay = d_opt, maxit = 1000)

train_pred_2 = predict(nn_2, train)
test_pred_2 = predict(nn_2, test)

#Neural network with 5 hidden neurons
nn_4 <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size = 4,
                decay = d_opt, maxit = 1000)

train_pred_4 = predict(nn_4, train)
test_pred_4 = predict(nn_4, test)

#Finding optimal size

err1 <- vector("numeric", 29)
s <- seq(2, 20, length.out = 29)
t = 1
for(i in s){
  nn_optsize <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, 
                      size = i, decay = d_opt, maxit = 1000)
  pred1.class <- predict(nn_optsize, newdata = valid)
  err1[t] <- sqrt(mean((pred.class-valid[,1])^2))
  t <- t+1
}

plot(s, err1)
min_index1 <- which.min(err1)
s_opt <- s[min_index1]

#Neural network with optimal size and decay
nn_opt <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size =
                 s_opt, decay = d_opt, maxit = 1000)

train_pred_opt = predict(nn_opt,train)
test_pred_opt = predict(nn_opt,test)

Accuracy_train(train_pred_opt)
Accuracy_test (test_pred_opt)

#Neural network model with 8 hidden neurons
nn_8 <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size = 8,
                decay = d_opt, maxit = 1000)

train_pred_8 = predict(nn_8, train)
test_pred_8 = predict(nn_8, test)

#Neural network model with 10 hidden neurons
nn_10 <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size = 10,
                decay = d_opt, maxit = 1000)

train_pred_10 = predict(nn_10, train)
test_pred_10 = predict(nn_10, test)

#Neural network model with 12 hidden neurons
nn_12 <- nnet(Price ~ HP + CC + Grs + Wght + AC + M_Rim, data = train, linout = FALSE, size = 12,
                decay = d_opt, maxit = 1000)

train_pred_12 = predict(nn_12, train)
test_pred_12 = predict(nn_12, test)

#Aggregating the plots
#plotnet(nn_opt)

```

Now, let's see the accuracy of the various models on train and test data
```{r}
table <- data.frame(
  size = c(2,4,8,10,12),
  train_accuracy = c(Accuracy_train(train_pred_2), Accuracy_train(train_pred_4), Accuracy_train(train_pred_8), Accuracy_train(train_pred_10), Accuracy_train(train_pred_12)),
  test_accuracy = c(Accuracy_test(test_pred_2), Accuracy_test(test_pred_4), Accuracy_test(test_pred_8), Accuracy_test(test_pred_10), Accuracy_test(test_pred_12))
)

table

max_acc_ind <- which.max(table[,3])
best_model_size <- table[max_acc_ind,1]
```


The `r best_model_size` hidden neuron NN model is generating highest accuracy with the accuracy on test data = `r table[max_acc_ind,3]`.

As the best model size is smaller, we can interpret that the model is less prone to over fitting to the train data. Since, the dataset we trained the model on is small, the complexiy of the model is lower. But, the accuracy on the unseen data is better with less complicated model because of good generalization capability of the lower size model.


### (3) **Compare your neural network models with linear regression model. Which one is better?**

**(A)** Let's create a Linear regression model for the dataset.

```{r lmdata}
cd_lm$Colour <- as.factor(cd_lm$Colour)
str(cd_lm)
normalize <- function(x) 
{
  (x-min(x))/(max(x)-min(x))
}

cleandata_lm <- cd_lm %>% mutate_if(is.numeric, normalize)
cleandata_lm <- cleandata_lm[,-6]

str(cleandata_lm)
set.seed(90)
indx <- sample(2, nrow(cleandata_lm), replace = T, prob = c(0.7,0.3))
lm_train <- cleandata_lm[indx == 1, ]
lm_test <- cleandata_lm[indx == 2, ]

```

Doing forward selection for the important variables and creating the linear regression model

```{r}
full <- lm(lm_train$Price ~., data = lm_train)
null <- lm(lm_train$Price ~1, data = lm_train)

# Forward Selection
forward <- step(null, scope = list(lower = null, upper = full), direction = "forward", trace = 0)

lmModel <- lm(formula(forward), data = lm_train)
plot(lmModel)
```

```{r pred_lm}
lmpred_train <- predict(lmModel, lm_train)
lmpred_test <- predict(lmModel, lm_test)

lmtrain_error <- Metrics::rmse(lmpred_train, lm_train$Price)
lmtest_error <- Metrics::rmse(lmpred_test, lm_test$Price)

lmtrain_error_percent <- lmtrain_error/mean(lm_train$Price)
lmtest_error_percent <- lmtest_error/mean(lm_test$Price)

```

Accuracy of linear regression model on test data = `r 1-lmtest_error_percent`

The accuracy of Neural network on the test data(`r table[max_acc_ind,3]`) is significantly higher than the accuracy of Linear regression model(`r 1-lmtest_error_percent`). Therefore, Neural network model is better for predicting car prices.


### (4) **Make a decision and offer your recommendations.**

**(A)** 
**Decision:** As the accuracy for neural network is significantly higher than the accuracy for linear regression model, Farid should choose the neural network model to predict the prices of the new manufactured cars. 

**Recommendation:** To increase accuracy of the models, a bigger dataset is required to train the models. For this dataset, as the number of provided instances were only 31, a small neural network was able to generalize the small dataset and predict the prices. But, predicting car prices is a complicated process and a bigger dataset would help the model better by increasing the complexity of the model(by increasing neurons and hiden layers) thereby increasing the accuracy.

In addition, the variables which matter the most to price would also change if there was a bigger dataset. As the KMs, Age didn't show any material correlation.
